Dark Control: The Default Mode Network as a Reinforcement Learning Agent

### 2018-08-07 paper/paper.tex
Danilo:additions for cortexrevision

This \textit{dark matter of brain physiology}
\citep{raichle2006brain} begs the question
of the biological purpose underlying DMN activity.
Despite observation of similar large-scale networks of co-varying spontaneous activity
in electrophysiological investigations \citep{de2010temporal,brookes2011investigating,baker2014fast}, the link between the fMRI BOLD signal and population-level neural activity is still unclear. If those frequency-specific electrophysiological correlations are proposed as complementary to those observed with BOLD \citep{hipp2015bold}, their role in DMN function remains elusive \citep{maldjian2014graph}.

What has early been described as the ``stream of consciousness''
in psychology \citep{james1890principles}
found a potential neurobiological manifestation
in the DMN
\citep{shul1997, raichle2001pnas}.
We propose that this set of some of the most advanced regions in
the association cortex \citep{mesulam1998sensation, margulies2016situating}
are responsible for higher-order control of human behavior.
Our functional account follows the notion of
``a hierarchy of brain systems with
the DMN at the top and the salience and dorsal attention systems
at intermediate levels, above thalamic and unimodal sensory
cortex'' \citep{carhart2010default}.

@@

\section{Known neurobiological properties of the default mode network}
\label{sec:nodes}
We begin by a neurobiological deconstruction of the DMN
based on experimental findings in the neuroscience literature.
This walkthrough across main regions of the DMN will
outline the individual functional profiles,
paving the way for their algorithmic interpretation
in our formal account (section 3).

\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
  \caption{\textbf{Default mode network: key functions.}
  Neurobiological overview of the DMN with its major constituent parts and
  the associated functional roles relevant in our functional interpretation.}
  \label{fig:neurobiological_overview_DMN}
\end{figure}
%
\subsection{The posteromedial cortex: global monitoring and information integration}
The midline structures of the human DMN,
including the posteromedial cortex (PMC) and
the medial prefrontal cortex (mPFC),
are probably responsible for highest turn-overs
of energy consumption \citep{raichle2001pnas, raichle_baseline}.
These metabolic characteristics go hand-in-hand with
brain-imaging findings that suggested the PMC and mPFC to
potentially represent the functional core of the DMN
\citep{andrews2010, hagmann2008mapping}.

Normal and disturbed metabolic fluctuations in the
human PMC have been closely related to
changes of conscious awareness \citep{cavanna2006precuneus}.
Indeed,
the PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication.
%
This DMN region has long been speculated to
reflect constant computation of
environmental statistics and its internal representation
as an inner ``mind's eye'' \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from medial damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Such neurological patients are plagued by an
inability to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as an inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This dysfunction can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two different functional connectivity analyses
\citep{bzdok2015subspecialization}
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological relevance assessment
\citep{amygdala_bzdok, ghods2009fundamental, baxter2002amygdala}.

@@

Consequently, the PMC may contribute to overall DMN function
by monitoring the subjective outcomes
of possible actions and integrating that information
with memory and perspective frames
into short- and longer-term behavioral agendas.
Estimated value, found to differs across individuals, might enrich
statistical assessment of the environment
to map and predict delayed reward opportunities in the future.
In doing so, the PMC may continuously adapt the organism to changes
in both the external environment and its internal representation
to enable strategic behavior.

\subsection{The prefrontal cortex: action consideration and stimulus-value association}
Analogous to the PMC,
the dorsomedial PFC (dmPFC) of the DMN is believed to subserve
multi-sensory processes
across time, space, and content domains to
exert top-level control on behavior.
Comparing to the PMC, however,
dmPFC function may be closer to a
``mental sketchpad'' \citep{goldman1996prefrontal}, as this DMN part
potentially subserves the de-novo construction and manipulation
of meaning representations instructed by stored semantics and memories
\citep{bzdok2013segregation}.
The dmPFC may subserve representation and assessment
of one's own and other individuals' action considerations.
Generally,
neurological patients with tissue damage in the prefrontal cortex
are known to struggle with
adaptation to new stimuli and events
\citep{stuss1986frontal}.
Specifically, neural activity in the human dmPFC
reflected expectations about other peoples' actions and
outcomes of these predictions.
Neural activity in the dmPFC indeed explained the performance decline
of inferring other peoples' thoughts in aging humans \citep{moran2012social}.
Certain dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, action
with fine-grained adjustment of contextual aspects \citep{yoshida2010neural}.
% In fact, the spatially neighboring dorsal anterior cingulate cortex
% has also been linked to computing values and efforts of
% persisting a behavioral plan versus switching the
% environmental context in several lesion studies \citep{kolling2016value}.
%
% Such highly abstract neural computations necessarily rely on the
% construction of probabilistic internal information drawing from
% memory recall, mental scene elaboration processes,
% and stored knowledge of world regularities.
% %
% Moreover,
% in a computational brain-imaging experiment,
% dorsomedial PFC activity preferentially processed the consequences of
% action choices that were considered but not actually executed,
% whereas ventromedial PFC (vmPFC) activity
% processed especially value outcomes of actions actually performed on the environment
% \citep{nicolle2012agent}.

@@

% \begin{wrapfigure}{r}{.4\linewidth}
\begin{figure}[!htbp]
  \begin{minipage}[c]{0.63\textwidth}
    \includegraphics[width=.8\linewidth]{fig_dmn.png}
  \end{minipage}
  \hspace{-4em}
  \begin{minipage}[c]{0.45\textwidth}
    \caption{\textbf{Morphological coupling between reward system and
        default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from volume in the DMN regions. These out-of-sample generalization performances were obtained from support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure.
      Consistent for the left and right reward system, NAc volume in a given subject is
      positively coupled with the vmPFC and HC. The colors are indicative of the (\textcolor{red}{red} = positive,
      \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
  The code for reproduction and visualization:
  \url{www.github.com/banilo/darkcontrol_2018}.
  }
  \label{fig:dmn}
\end{minipage}
% \end{wrapfigure}
\end{figure}

@@

\subsection{The hippocampus: memory, space, and experience replay}
The DMN midline has close functional links
with the HC in the medial temporal lobe \citep{vincet2006, shannon2013morning} \textemdash
a region long known to be involved in
memory operations and spatial navigation in animals and humans.
While the HC
is traditionally believed to allow recalling past experience,
there is now increasing evidence for an important role
in constructing mental models in general
\citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017,
boyer2008evolutionary}.
Its recursive anatomical architecture
may be specifically designed to allow reconstructing
entire sequences of experience from memory fragments.
Indeed,
hippocampal damage was
not only associated with an impairment in re-experiencing the past (i.e., amnesia),
but also forecasting of one's own future and
imagination of experiences more broadly \citep{hassabis2007patients}.

@@

There is hence increasing evidence that
HC function extends beyond simple forms of
encoding and reconstruction of memory and space information.
Based on spike recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events \citep{buzsaki2004large}.
Specific spiking sequences, which were elicited by experimental task design,
have been shown to be re-enacted spontaneously during
quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
Moreover, neuronal spike sequences measured in hippocampal place cells of rats
featured re-occurrence directly after experimental trials
as well as directly before (prediction of) upcoming experimental trials \citep{diba2007forward}.
Similar spiking patterns in hippocampal neurons during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage, potentially including DMN regions.
Moreover, in mice, invasively triggering spatial experience recall
in the HC during sleep
has been demonstrated to subsequently alter action choice during wakefulness
\citep{lavilleon2015}.
These HC-subserved mechanisms
conceivably contribute to advanced cognitive processes that require
re-experiencing or newly constructed mental scenarios,
such as in recalling autobiographical memory episodes
\citep{hassabis2007patients}.
Thus, the HC would orchestrate re-experience of environmental aspects for
consolidations based on re-enactment and for integration into
rich mental scene construction \citep{deuker2016event, bird2010establishing}.
As such, the HC may impact
ongoing perception of and action on the environment
\citep{maguire2016, lavilleon2015}.
% (TODO: relate to self-play in deepmind Go).
% (TODO: Kai had an interesting idea: the deja-vu effect
% could be when you already accidentally perturbed past
% such that it really occures in reality later.)

@@

The TPJ in the right hemisphere (RTPJ)
has been shown to be closely related to
multi-sensory prediction and prediction error signaling.
This DMN region is probably central for
action initiation during goal-directed psychological tasks and for
sensorimotor behavior by integrating multi-sensory attention
\citep{corbetta2002control}.
Its involvement was repeatedly reported in
multi-step action execution \citep{hartmann2005takes},
visuo-proprioceptive conflict \citep{Balslev2005}, and
detection of environmental changes across
visual, auditory, or tactile stimulation
\citep{downar2000multimodal}.
Direct electrical stimulation of the human
RTPJ during neurosurgery was associated with altered perception
and stimulus awareness \citep{blanke2002neuropsychology}.
%
It was argued that the RTPJ encodes actions and predicted outcomes,
without necessarily relating these neural processes to value estimation
\citep{liljeholm2013neural, hamilton2008action,
jakobs2009effects}.
Additionally, neural activity in the RTPJ has been proposed to reflect
stimulus-driven attentional reallocation to
self-relevant and unexpected sources of information
as a ‘‘circuit breaker’’ that recalibrates functional control of brain networks
\citep{bzdok2013tpj, corbettashul2008}.
% (Bzdok et al., 2013; Corbetta et al., 2008).
Indeed, neurological patients with RTPJ damage have particular difficulties
with multi-step actions \citep{hartmann2005takes}.
In the face of large discrepancies between actual and previously predicted
environmental events, the RTPJ acts as a potential switch between
externally-oriented mind sets focussed on the
sensory environment and internally-oriented mind sets focussed
on mental scene construction.
For instance, temporally induced RTPJ damage in humans diminished the
impact of predicted intentions of other individuals
\citep{young2010disruption},
a capacity believed to be enabled by the DMN.
The RTPJ might hence be an important relay that shifts away
from the ‘‘internally directed’’ baseline processes
to, instead, deal with unexpected environmental stimuli and events.
% (TODO: slow endogeneous changes in state transitions by the policy and value matrices)

The left TPJ of the DMN (LTPJ),
in turn, has a close relationship to Wernicke's area
involved in semantic processes, such as in spoken and written language.
Neurological patients with damage in Wernicke's area
have a major impairment of language comprehension
when listening to others or reading a book.
Patient speech
preserves natural rhythm and normal syntax, yet the
voiced sentences lack meaning (i.e., aphasia).
Abstracting from speech interpretations in linguistics
and neuropsychology,
the LTPJ appears to mediate access to and binding of world knowledge,
such as required during action considerations
\citep{binder2011neurobiology, seghier2013angular}.
Consistent with this view,
LTPJ damage in humans also entailed problems in recognizing
others' pantomimed action towards objects
without obvious relation to processing explicit language content
\citep{varney1987locus}.
% (Varney and Damasio 1987; Rothi et al. 1991; Buxbaum et al. 2005).
%
Inner speech also hinges on knowledge recall
about the physical and social world.
Indeed,
the internal production of
verbalized thought (``language of the mind'') was closely related to the LTPJ
in a pattern analysis of brain volume
\citep{geva2011neural}.
Further,
episodic memory recall and mental imagery to forecast future events strongly draw on
re-assembling world knowledge.
Isolated building blocks of world structure get rebuilt
in internally constructed mental scenarios that
guide present action choice,
weigh hypothetical possibilities, and forecast event outcomes.
%
Neural processes in the LTPJ may hence
contribute to the automated predictions of the environment
by incorporating experience-derived building blocks of world regularities
into ongoing action, planning, and problem solving.
% world -> policy matrix driven computation
% self -> value matrix driven computation

@@

\subsection{Putting everything together}
The DMN is today known to consistently increase in neural
activity when humans engage in cognitive processes that are relatively detached from
the current sensory environment.
The more familiar and predictable the current environment, the more brain resources may remain for allocating DMN activity to MDP processes extending beyond the present time and sensory context. In line with this perspective, DMN engagement was shown to heighten and relate to effective behavioral responses in the practiced phase of a demanding cognitive flexibility task, as compared to acquisition phase when participants learned context-specific rules. This involvement in automated decision-making has led the authors to propose an ``autopilot'' role for the DMN \citep{vatansever2017pnas}, which may contribute to optimizing intervention of the organism on the world in general. Among all parts of the DMN, the RTPJ is perhaps the most evident candidate for a network-switching relay that calibrates between processing of environment-engaged versus internally generated information \citep{downar2000multimodal, golland2006extrinsic, bzdok2013tpj}.

Additionally,
the DMN was proposed to be situated
at the top of the brain network hierarchy, with
the subordinate salience and dorsal attention network in the middle and
the primary sensory cortices at the bottom
\citep{carhart2010default, margulies2016situating}.
Its putative involvement in thinking about 
hypothetical experiences and future outcomes
appears to tie in with the implicit computation of
action and state cascades as a function of
experienced events and collected feedback from the past.
A policy matrix encapsulates the choice probabilities of possible actions
on the world given a current situation (i.e., state).
The DMN may subserve
  constant exploration of candidate action trajectories and
  nested estimaton of their
  cumulative reward outcomes. Implicit computation of future choices
  provides a potential explanation for the
  evolutionary emergence and practical usefulness of
  mind-wandering at day-time and dreams during sleep
  in humans.
\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
  \caption{\textbf{Default mode network:
  possible neurobiological implementation of reinforcement learning.}
  Overview of how the constituent regions of the DMN (refer to section \ref{sec:nodes}) may
  map onto computational components necessary for an RL agent.}
  \label{fig:rl_process_chart}
\end{figure}

>>>>

This \textit{dark matter of brain physiology}
\citep{raichle2006brain} begs the question
of the biological purpose underlying DMN activity,
which however still remains elusive at the electrophysiological level
\citep{de2010temporal,brookes2011investigating,baker2014fast}.
% the link between the fMRI BOLD signal and population-level neural activity is still unclear. If those frequency-specific electrophysiological correlations are proposed as complementary to those observed with BOLD \citep{hipp2015bold}, their role in DMN function remains elusive \citep{maldjian2014graph}.
What has early been described as the ``stream of consciousness''
in psychology \citep{james1890principles}
found a potential neurobiological manifestation
in the DMN
\citep{shul1997, raichle2001pnas}.
%
Axonal tracing injection in such parts of the association cortex in monkeys
were shown to resemble connectivity links between nodes of the human
DMN (see here for details on anatomical connections: \citep{randy2008}).
Additionally, myelination patterns of axon connections were found to finish
particularly late in these cortical areas \citep{flechsig1920}, often believed to suggest
sophistication of subserved neural processes
\citep{Sowell2003, Yakovlev1967}.
%
We propose that this set of some of the most advanced regions in
the association cortex \citep{mesulam1998sensation, margulies2016situating}
are responsible for higher-order control of human behavior.
Our functional account follows the notion of
``a hierarchy of brain systems with
the DMN at the top and the salience and dorsal attention systems
at intermediate levels, above thalamic and unimodal sensory
cortex'' \citep{carhart2010default}.

@@

\section{Known neurobiological properties of the default mode network}
\label{sec:nodes}
We begin by a neurobiological deconstruction of the DMN
based on integrating experimental findings in the neuroscience literature
from different species.
This walkthrough across main functional zones of the DMN
(i.e., de-emphasizing their precise anatomical properties)
will outline the individual functional profiles with the goal of
paving the way for their algorithmic interpretation
in our formal account (section 3).
As our focus will be on major \textit{functional} zones of the DMN,
please see elsewhere for excellent surveys on their \textit{anatomical}
boundaries and connectivity patterns \citep{randy2008, binder2009, seghier2013angular}.

\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
  \caption{\textbf{Default mode network: key functions.}
  Neurobiological overview of the DMN with its major constituent parts and
  the associated functional roles relevant in our functional interpretation.
  The blue horizontal line indicates the cytoarchitectonic border between the
  more dorsal BA9 and the more ventral BA10 (Brodmann, 1909).
  %
  Axonal tracing in monkeys and diffusion tractography in humans suggested
  that the NAc of the reward circuitry has monosynaptic fiber connections
  to the vmPFC \citep{haber1995, croxson2005quantitative}. Evaluation of
  propagated value information and triggered affective states encoded
  in the vmPFC may then feed into the functionally connected partner nodes of
  the DMN, such as the dmPFC and PMC \citep{andrews2010, bzdok2013segregation}.
  }
  \label{fig:neurobiological_overview_DMN}
\end{figure}
%
\subsection{The posteromedial cortex: global monitoring and information integration}
The midline structures of the human DMN,
including the posteromedial cortex (PMC) and
the medial prefrontal cortex (mPFC),
are probably responsible for highest turn-overs
of energy consumption \citep{raichle2001pnas, raichle_baseline}.
These metabolic characteristics go hand-in-hand with
brain-imaging findings that suggested the PMC and mPFC to
potentially represent the functional core of the DMN
\citep{andrews2010, hagmann2008mapping}.

Normal and disturbed metabolic fluctuations in the
human PMC have been closely related to
changes of conscious awareness \citep{cavanna2006precuneus}.
Indeed,
the PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication.
%
This DMN region has long been speculated to
reflect constant computation of
environmental statistics and its internal representation
as an inner ``mind's eye'' \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that can result from tissue damage in the posterior medial cortex
\citep{balint1909seelenlahmung, randy2008}.
Such neurological patients are plagued by an
inability to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as an inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
%
Scanning complex scenes is impaired in that statistic or moving objects in the
environment may be invisible or disappear in the subject perception of the
patient \citep{mesulam_book, blumenfeld_book}.
%
This dysfunction can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
using these environmental opportunities towards a behavioral goal (i.e., exploitation).
Congruently,
the human PMC was coupled in two different functional connectivity analyses
\citep{bzdok2015subspecialization}
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological relevance assessment
\citep{amygdala_bzdok, ghods2009fundamental, baxter2002amygdala}.

@@

Consequently, the PMC may contribute to overall DMN function
by monitoring the subjective outcomes
of possible actions and integrating that information
with memory and perspective frames
into short- and longer-term behavioral agendas.
Estimated value, found to differs across individuals, might enrich
statistical assessment of the environment
to map and predict delayed reward opportunities in the future.
Viewed from a RL perspective, the PMC may continuously adapt the organism to changes
in both the external environment and its internal representation
to enable strategic behavior.

\subsection{The prefrontal cortex: action consideration and stimulus-value association}
Analogous to the PMC,
the dorsomedial PFC (dmPFC) of the DMN is believed to subserve
multi-sensory processes
across time, space, and content domains to
exert top-level control on behavior.
Comparing to the PMC, however,
dmPFC function may be closer to a
``mental sketchpad'' \citep{goldman1996prefrontal}, as this DMN part
potentially subserves the de-novo construction and manipulation
of meaning representations instructed by stored semantics and memories
\citep{bzdok2013segregation}.
The dmPFC may subserve representation and assessment
of one's own and other individuals' action considerations - a necessary
component of a full-blown RL agent.
Generally,
neurological patients with tissue damage in the prefrontal cortex
are known to struggle with
adaptation to new stimuli and events
\citep{stuss1986frontal}.
Specifically, neural activity in the human dmPFC
reflected expectations about other peoples' actions and
outcomes of these predictions.
Neural activity in the dmPFC indeed explained the performance decline
of inferring other peoples' thoughts in aging humans \citep{moran2012social}.
Certain dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, action
with fine-grained adjustment of contextual aspects \citep{yoshida2010neural}.
% In fact, the spatially neighboring dorsal anterior cingulate cortex
% has also been linked to computing values and efforts of
% persisting a behavioral plan versus switching the
% environmental context in several lesion studies \citep{kolling2016value}.
%
% Such highly abstract neural computations necessarily rely on the
% construction of probabilistic internal information drawing from
% memory recall, mental scene elaboration processes,
% and stored knowledge of world regularities.
% %
% Moreover,
% in a computational brain-imaging experiment,
% dorsomedial PFC activity preferentially processed the consequences of
% action choices that were considered but not actually executed,
% whereas ventromedial PFC (vmPFC) activity
% processed especially value outcomes of actions actually performed on the environment
% \citep{nicolle2012agent}.

@@

% \begin{wrapfigure}{r}{.4\linewidth}
\begin{figure}[!htbp]
  \begin{minipage}[c]{0.63\textwidth}
    \includegraphics[width=.8\linewidth]{fig_dmn.png}
  \end{minipage}
  \hspace{-4em}
  \begin{minipage}[c]{0.45\textwidth}
    \caption{\textbf{Predictive structural association between reward system and
        default mode network.} Reward tasks \citep{doherty2015structure} and neural processign in the DMN \citep{randy2008}, often call task-negative, have been studied so far in largely separate niches of the neuroscience literature. A currently underappreciated link is however suggested here based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$ [standard deviation across cross-validation folds]) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from (z-scored) volume in the DMN regions. These out-of-sample generalizations reflect the expected performance in yet-to-be observed individuals obtained from linear support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure \citep{hastie}.
      Consistent for the left and right reward system, NAc volume in a given subject is
      positively coupled with the vmPFC and HC. The congruence of our structural association results for both NAc targets speaks to the robustness of our pattern-prediction findings. The opposite relation of the left and right TPJ to the NAc appears to reflect a repeatedly recognized hemispheric asymmetry with respect to functional implications \citep{seghier2013angular}, impairments in neurological patients \citep{corbetta2000}, different types of connectivity \citep{uddin_tpj, caspers2011} as well as micro- and macroanatomy \citep{caspers2006, caspers2008}. The colors are indicative of the (\textcolor{red}{red} = positive,
      \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
  The code for reproduction and visualization:
  \url{www.github.com/banilo/darkcontrol_2018}.
  }
  \label{fig:dmn}
\end{minipage}
% \end{wrapfigure}
\end{figure}

@@

\subsection{The hippocampus: memory, space, and experience replay}
The DMN midline has close functional links
with the HC \citep{henceforth implying to include also parahippocampal regions}
in the medial temporal lobe \citep{vincet2006, shannon2013morning} \textemdash
a region long known to be involved in
memory operations and spatial navigation in animals and humans.
While the HC
is traditionally believed to allow recalling past experience,
there is now increasing evidence for an important role
in constructing mental models in general
\citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017,
boyer2008evolutionary}.
Its recursive anatomical architecture
may be specifically designed to allow reconstructing
entire sequences of experience from memory fragments.
Indeed,
hippocampal damage was
not only associated with an impairment in re-experiencing the past (i.e., amnesia),
but also forecasting of one's own future and
imagination of experiences more broadly \citep{hassabis2007patients}.

@@

There is hence increasing evidence that
HC function extends beyond simple forms of
encoding and reconstruction of memory and space information.
Based on spike recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events \citep{buzsaki2004large}.
Specific spiking sequences, which were elicited by experimental task design,
have been shown to be re-enacted spontaneously during
quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
Moreover, neuronal spike sequences measured in hippocampal place cells of rats
featured re-occurrence directly after experimental trials
as well as directly before (prediction of) upcoming experimental trials \citep{diba2007forward}.
Similar spiking patterns in hippocampal neurons during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage, potentially including DMN regions.
Moreover, in mice, invasively triggering spatial experience recall
in the HC during sleep
has been demonstrated to subsequently alter action choice during wakefulness
\citep{lavilleon2015}.
These HC-subserved mechanisms
conceivably contribute to advanced cognitive processes that require
re-experiencing or newly constructed mental scenarios,
such as in recalling autobiographical memory episodes
\citep{hassabis2007patients}.
Within a RL framework, the HC could thus orchestrate re-experience of environmental aspects for
consolidations based on re-enactment and for integration into
rich mental scene construction \citep{deuker2016event, bird2010establishing}.
As such, the HC may impact
ongoing perception of and action on the environment
\citep{maguire2016, lavilleon2015}.
% (TODO: relate to self-play in deepmind Go).
% (TODO: Kai had an interesting idea: the deja-vu effect
% could be when you already accidentally perturbed past
% such that it really occures in reality later.)

@@

The TPJ in the right hemisphere (RTPJ) denotes a broad functional zone with
varying anatomical nomenclature \citep{marstpj2011, seghier2010, seghier2013angular}
that has been shown to be closely related to
multi-sensory prediction and prediction error signaling
\citep{downar2000multimodal, vetter2011, shulman2010, shulman2007}.
This DMN region is probably central for
action initiation during goal-directed psychological tasks and for
sensorimotor behavior by integrating multi-sensory attention
\citep{corbetta2002control}.
Its involvement was repeatedly reported in
multi-step action execution \citep{hartmann2005takes},
visuo-proprioceptive conflict \citep{Balslev2005}, and
detection of environmental changes across
visual, auditory, or tactile stimulation
\citep{downar2000multimodal}.
Direct electrical stimulation of the human
RTPJ during neurosurgery was associated with altered perception
and stimulus awareness \citep{blanke2002neuropsychology}.
%
It was argued that the RTPJ encodes actions and predicted outcomes,
without necessarily relating these neural processes to value estimation
\citep{liljeholm2013neural, hamilton2008action,
jakobs2009effects}.
Additionally, neural activity in the RTPJ has been proposed to reflect
stimulus-driven attentional reallocation to
self-relevant and unexpected sources of information
as a ‘‘circuit breaker’’ that recalibrates functional control of brain networks
\citep{bzdok2013tpj, corbettashul2008}.
% (Bzdok et al., 2013; Corbetta et al., 2008).
Indeed, neurological patients with RTPJ damage have particular difficulties
with multi-step actions \citep{hartmann2005takes}.
In the face of large discrepancies between actual and previously predicted
environmental events, the RTPJ acts as a potential switch between
externally-oriented mind sets focussed on the
sensory environment and internally-oriented mind sets focussed
on mental scene construction.
For instance, temporally induced RTPJ damage in humans diminished the
impact of predicted intentions of other individuals
\citep{young2010disruption},
a capacity believed to be enabled by the DMN.
Viewed from a RL account,
the RTPJ might reflect an important relay that shifts away
from the ‘‘internally directed’’ baseline processes
to, instead, deal with unexpected environmental stimuli and events.
% (TODO: slow endogeneous changes in state transitions by the policy and value matrices)

The left TPJ of the DMN (LTPJ),
in turn, may have a functional relationship to Wernicke's area
involved in semantic processes \citep{blumenfeld_book} and has
been described as ``a temporoparietal transmodal gateway for language''
by some investigators \citep{mesulam_book}.
Neurological patients with damage in Wernicke's area
have a major impairment of language comprehension
when listening to others or reading a book.
Patient speech
preserves natural rhythm and normal syntax, yet the
voiced sentences lack meaning (i.e., aphasia).
Abstracting from speech interpretations in linguistics
and neuropsychology,
the LTPJ appears to mediate access to and binding of world knowledge,
such as required during action considerations
\citep{binder2011neurobiology, seghier2013angular}.
Consistent with this view,
LTPJ damage in humans also entailed problems in recognizing
others' pantomimed action towards objects
without obvious relation to processing explicit language content
\citep{varney1987locus}.
% (Varney and Damasio 1987; Rothi et al. 1991; Buxbaum et al. 2005).
%
Inner speech also hinges on knowledge recall
about the physical and social world.
Indeed,
the internal production of
verbalized thought (``language of the mind'') was closely related to the LTPJ
in a pattern analysis of brain volume
\citep{geva2011neural}.
Further,
episodic memory recall and mental imagery to forecast future events strongly draw on
re-assembling world knowledge.
Isolated building blocks of world structure get rebuilt
in internally constructed mental scenarios that
guide present action choice,
weigh hypothetical possibilities, and forecast event outcomes.
%
As a candidate component of a RL agent,
neural processes in the LTPJ may
contribute to the automated predictions of the environment
by incorporating experience-derived building blocks of world regularities
into ongoing action, planning, and problem solving.
% world -> policy matrix driven computation
% self -> value matrix driven computation

@@

\subsection{Putting everything together}
The DMN is today known to consistently increase in neural
activity when humans engage in cognitive processes that are relatively detached from
the current sensory environment.
The more familiar and predictable the current environment, the more brain resources may remain for allocating DMN activity to MDP processes extending beyond the present time and sensory context.
%
This speculation receives quantitative support in that connectional links between
nodes of the DMN have been reported to be more consistent and reliable than
functional couplings within any other macroscopical networks \citep{shehz2009}.
As such, random-sampling-related baseline evaluation of possibilities and
events may be subserved by the DMN and get partly suspended when novelty in
the external environment is encountered or immediate action is required.
%
In line with this perspective, DMN engagement was shown to heighten and relate to effective behavioral responses in the practiced phase of a demanding cognitive flexibility task, as compared to acquisition phase when participants learned context-specific rules. This involvement in automated decision-making has led the authors to propose an ``autopilot'' role for the DMN \citep{vatansever2017pnas}, which may contribute to optimizing intervention of the organism on the world in general. Among all parts of the DMN, the RTPJ is perhaps the most evident candidate for a network-switching relay that calibrates between processing of environment-engaged versus internally generated information \citep{downar2000multimodal, golland2006extrinsic, bzdok2013tpj}.

Additionally,
the DMN was proposed to be situated
at the top of the brain network hierarchy, with
the subordinate salience and dorsal attention network in the middle and
the primary sensory cortices at the bottom
\citep{carhart2010default, margulies2016situating}.
Its putative involvement in thinking about 
hypothetical experiences and future outcomes
appears to tie in with the implicit computation of
action and state cascades as a function of
experienced events and collected feedback from the past.
A policy matrix encapsulates the choice probabilities of possible actions
on the world given a current situation (i.e., state).
The DMN may subserve
  constant exploration of candidate action trajectories and
  nested estimaton of their
  cumulative reward outcomes. Implicit computation of future choices
  provides a potential explanation for the
  evolutionary emergence and practical usefulness of
  mind-wandering at day-time and dreams during sleep
  in humans.
\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
  \caption{\textbf{Default mode network:
  possible neurobiological implementation of reinforcement learning.}
  Overview of how the constituent regions of the DMN (refer to section \ref{sec:nodes}) may
  map onto computational components necessary for an RL agent.
  %
  Axonal tracing in monkeys and diffusion tractography in humans suggested
  that the NAc of the reward circuitry has monosynaptic fiber connections
  to the vmPFC \citep{haber1995, croxson2005quantitative}. Evaluation of
  propagated value information and triggered affective states encoded
  in the vmPFC may then feed into the functionally connected partner nodes of
  the DMN, such as the dmPFC and PMC \citep{andrews2010, bzdok2013segregation}.
}
  \label{fig:rl_process_chart}
\end{figure}
